%&pdflatex 
\documentclass[12pt]{article} 
\usepackage[margin=1in]{geometry} 
\usepackage{amsmath,amsthm,amssymb,amsfonts,tikz-cd} 

\newenvironment{problem}[2][Problem]{\begin{trivlist}
\item[\hskip \labelsep {\bfseries #1}\hskip \labelsep {\bfseries #2.}]}{\end{trivlist}}

\newenvironment{proposition}[1][Proposition]{\begin{trivlist}
\item[\hskip \labelsep {\bfseries #1.}]}{\end{trivlist}}

\newenvironment{definition}[1][Definition]{\begin{trivlist}
\item[\hskip \labelsep {\bfseries #1.}]}{\end{trivlist}}

\newcommand{\til}{\char`\~}
\newcommand{\bs}{\textbackslash} 

\newcommand{\catname}[1]{\normalfont\textbf{#1}}
\newcommand{\catsup}[2]{\normalfont\textbf{#1}^{#2}}
\newcommand{\catsub}[2]{\normalfont\textbf{#1}_{#2}}


\newcommand{\Hom}{\text{Hom}}
\newcommand{\Homc}[2]{\Hom_{\catname{#1}}(#2)}

\newcommand{\Obj}{\text{Obj}}
\newcommand{\Objc}[1]{\text{Obj}(\catname{#1})}
\newcommand{\Aut}{\text{Aut}}
\newcommand{\End}{\text{End}}

\newcommand{\id}{\text{id}}

\newenvironment{solution}
  {\renewcommand\qedsymbol{$\blacksquare$}\begin{proof}[Solution]}
{\end{proof}}

\newenvironment{sproof}{
  \renewcommand\qedsymbol{$\square$}
  \begin{proof}
  }{
  \end{proof}
}

\begin{document}

\title{Algebra: Chapter 0 Exercises\\ \large Chapter 1, Section 5}
\author{David Melendez}
\maketitle

\begin{problem}{5.1}
  A final object in a category \catname{C} is initial in the opposite category $\catname{C}^{op}$.
\end{problem}
\begin{proof}
  Let $T \in \Objc{C}$ be initial and $Z \in \Objc{C}$ be any object. 
  Since $\Homc{C}{T, Z}$ is a singleton, 
  $\Hom_{\catname{C}^{op}}(Z, T) = \Homc{C}{T, Z}$ 
  is also a singleton, making $T$ final in $\catname{C}^{op}$
\end{proof}

\begin{problem}{5.2}
  The empty set $\emptyset$ is the \textit{unique} initial object in \catname{Set}.
\end{problem}
\begin{proof}
  Recall that the number of morphisms between any two sets $A$ and $B$ is given by $B^A$. 
  Thus, the problem of finding the size of an initial object in \catname{Set} boils down to solving $|Z|^{|I|}=1$ for $|I|$, where $Z$ is an arbitrary set.
  The only solution to this is $|I|=0$, which is only satisfied by the the null set, given that $\emptyset^\emptyset = 0$.
\end{proof}

\begin{problem}{5.3}
  Final objects are unique up to isomorphism.
\end{problem}
\begin{proof}
  Let \catname{C} be a category and $T_1, T_2\in \Objc{C}$ be final. Because $T_1$ and $T_2$ are final, we have, for all $Z\in\Homc{C}{Z, T}$:
  \begin{align*}
    \End(T_1) &= \{\id_{T_1}\}\\
    \End(T_2) &= \{\id_{T_2}\}\\
    \Hom(T_1, T_2) &= \{f\} \text{ for some } f\\
    \Hom(T_2, T_1) &= \{g\} \text{ for some } g\\
  \end{align*}
  We then have $fg$ = $\id_{T_2}$ and $gf$ = $\id_{T_2}$, giving us $T_1 \cong T_2$.
\end{proof}

\newpage

\begin{problem}{5.4}
  What are the initial and final objects in the category of 'pointed sets' (Example 3.8)? Are they unique?
\end{problem}
\begin{solution}
  For the sake of completeness, a description of this category $\catsup{Set}{*}$ will be included.
  \begin{definition}
    Define $\catsup{Set}{*}$ as follows:\\
    Objects in $\catsup{Set}{*}$ are morphisms $f : \{*\} \to A$ where $A$ is any set, denoted $(f, A)$. \\
    Morphisms $\sigma \in \Hom_{\catsup{Set}{*}}( (f_A, A), (f_B, B))$ are commutative diagrams: 
    \[\begin{tikzcd}
        A \ar{rr}{\sigma} && B \\
        & \{*\} \ar{ul}{f_A} \ar{ur}[swap]{f_B}
    \end{tikzcd}\]
  \end{definition}
  Now, we will answer the question.
  \begin{proposition}
    Initial and final objects in $\catsup{Set}*$ are objects $(i, T)$, where $T$ is a singleton and $i$ is the unique function that maps from $\{*\}$ to $T$.
  \end{proposition}
  \begin{sproof}
    We will prove that the objects described are both initial and final in $\catsup{Set}*$.
    \begin{enumerate}
      \item Initial: \\
        Consider the object described above. 
        Then, a morphism from this object to another object 
        $(f, B)\in \Obj(\catsup{Set}*)$
        is given by the following commutative diagram:
        \[\begin{tikzcd}
          T \ar{rr}{\sigma} && B \\
          & \{*\} \ar{ul}{i} \ar{ur}[swap]{f}
        \end{tikzcd}\]
        The only choice of $\sigma$ that makes this diagram commute is defined by $\sigma(i(*)) = f(*)$ (as $T$ is a singleton), so $T$ is initial.

      \item Final: 
        Consider the object described above. 
        Then, a morphism to this object from another object 
        $(f, B)\in \Obj(\catsup{Set}*)$
        is given by the following commutative diagram:
        \[\begin{tikzcd}
          A \ar{rr}{\tau} && T \\
          & \{*\} \ar{ul}{f} \ar{ur}[swap]{i}
        \end{tikzcd}\]
        The only choice of $\tau$ that makes this diagram commute is defined by $\tau(f(*)) = i(*)$ (as $T$ is a singleton), so $T$ is final.
    \end{enumerate}
  \end{sproof}
\end{solution}

\newpage

\begin{problem}{5.5}
  What are the final objects in the category considered in 5.3?
\end{problem}
\begin{solution}
  Final objects in this category (let's denote it \catname{C}) are singletons.
  \begin{sproof}
    Let $\{*\}$ be a singleton.
    Then morphisms $\tau$ from $(Z, \varphi) \in \Objc{C}$ to $\{*\}$ are commutative diagrams:
    \[\begin{tikzcd}
        Z \ar{rr}{\tau} && \{*\} \\
        & A \ar{ul}{\varphi} \ar{ur}[swap]{i}
    \end{tikzcd}\]
    Since $\{*\}$ is a singleton, the only $\tau$ that satisfies
    $\tau(\varphi(a)) = \iota(a)$ is defined by $\tau(z) = *$ for all $z\in Z$.
    Thus $(\{*\}, \tau)$ is final in \catname{C}.
  \end{sproof}
\end{solution}

\begin{problem}{5.6}
  Consider the category corresponding to endowing (as in Example 3.3) the set $\mathbb{Z}^+$ with the divisibility relation. 
  Thus there is exactly one morphism $d \to m$ in this category iff $d$ divides $m$ without remainder; there is no morphism between $d$ and $m$ otherwise.
  Show that this category has products and coproducts. What are their 'conventional' names?
\end{problem}
\begin{solution}
  Let \catname{Div} denote this category and $a, b \in \Objc{Div}$ be objects.
  \begin{enumerate}
    \item Products\\
        That \catname{Div} has products means the following:\\
    For every $a, b \in \Objc{Div}$, 
    there exists an object $a\times b \in \Objc{Div}$ such that
    every $(x,f_a,f_b) \in \Obj(\catsub{Div}{a,b})$
    admits a unique morphism $\sigma \in \Homc{Div}{x, a\times b}$
    that makes the following diagram commute:
    \[\begin{tikzcd}
      &&&a\\
      &x \ar[bend left]{urr}{f_a} \ar{r}{\sigma} \ar[bend right]{drr}[swap]{f_b}
      &a\times b \ar{ur}{\pi_a} \ar{dr}[swap]{\pi_b}&\\
      &&&b
    \end{tikzcd}\]
    Meaning, for every pair of integers $(a,b)$, there exists a unique
    factor of a and b, $a\times b$, such that 
    every common factor of $a$ and $b$ divides $a\times b$.
    This is necessarily the greatest common factor of $a$ and $b$, since
    it must be a multiple of every common factor of $a$ and $b$, and 
    $\text{gcf}(a,b)$ is the product of the common factors of $a$ and $b$.
  \item Coproducts\\
    That \catname{Div} has coproducts means the following:\\
    For every $a, b \in \Objc{Div}$, 
    there exists an object $a\sqcup b \in \Objc{Div}$ such that
    every $(x,f_a,f_b) \in \Obj(\catsup{Div}{a,b})$
    admits a unique morphism $\sigma \in \Homc{Div}{a\sqcup b, x}$
    that makes the following diagram commute:
    \[\begin{tikzcd}
        &&&a \ar[bend right, swap]{dll}{\pi_a} \ar[swap]{dl}{f_a}\\
      &a\sqcup b \ar{r}{\sigma} 
      &x &\\
      &&&b \ar[bend left]{ull}{\pi_b} \ar{ul}{f_b}
    \end{tikzcd}\]
    Meaning, for every pair of integers $(a,b)$ there exists a unique multiple of $a$ and $b$, that divides every common multiple of $a$ and $b$. 
    Since $a\sqcup b$ is a common multiple of $a$ and $b$ and $a\sqcup b \leq x$ for every common multiple $x$ of $a$ and $b$, $a\sqcup b$ is $\text{lcm}(a,b)$.
  \end{enumerate}
\end{solution}

\begin{problem}{5.9}
  Let \catname{C} be a category with products. Find a reasonable candidate for the universal property that the product 
  $A\times B\times C$ 
  of \textit{three} objects of \catname{C} ought to satisfy, and prove that both 
  $(A\times B)\times C$ and
  $A\times (B\times C)$ 
  satisfy this universal property.
\end{problem}
\begin{solution}
  For objects $A,B,C$ of a category \catname{C}, $A\times B\times C$ ought to be universal with respect to the property of mapping to $A, B,$ and $C$.
  In other words, every $(X, f_a, f_b, f_c)$ in the category $\catsup{C}{A,B,C}$ admits a unique morphism $\sigma \in \Homc{C}{X, A\times B\times C}$ such that the following diagram commutes:\\
  \[\begin{tikzcd}{(1)}
    &&&A\\
    &X 
    \ar{r}{\sigma_{ABC}} 
    \ar[bend left]{urr}{f_A} 
    \ar[swap]{dr}{f_B}
    \ar[bend right, swap]{ddrr}{f_C}
    &A\times B\times C
    \ar[swap]{ur}{\pi_A}
    \ar{d}{\pi_B}
    \ar[bend left]{ddr}{\pi_C}
    &\\
    &&B&\\
    &&&C
  \end{tikzcd}\]
Meaning, $\pi_A \sigma = f_A$, $\pi_B \sigma = f_B$, and $\pi_C \sigma = f_C$. \\
\begin{proposition}
  The products $A\times (B\times C)$ and $(A\times B)\times C$ satisfy this property.
\end{proposition}
\begin{proof}
  Consider the product $A\times (B\times C)\in \Objc{C}$ and its diagram:
  \[\begin{tikzcd}{(2)}
    &&&A\\
    &X 
    \ar{r}{\sigma_{A(BC)}} 
    \ar[bend left]{urr}{f_A} 
    \ar[bend right, swap]{drr}{f_{BC}}
    &A\times (B\times C)
    \ar[swap]{ur}{\phi_A}
    \ar{dr}{\phi_{BC}}
    &\\
    &&&B\times C
  \end{tikzcd}\]
  along with the product $B\times C$ and its diagram:
  \[\begin{tikzcd}{(3)}
    &&&B\\
    &X 
    \ar{r}{\sigma_{BC}} 
    \ar[bend left]{urr}{f_B} 
    \ar[bend right, swap]{drr}{f_C}
    &B\times C
    \ar[swap]{ur}{\phi_B}
    \ar{dr}{\phi_C}
    &\\
    &&&C
  \end{tikzcd}\]
  Staring at these diagrams,
  we can see that 
  $A\times (B\times C)$ satisfies our universal property if we let:
  $f_{BC} = \sigma_{BC}$,\\
  $\pi_A = \phi_A$,\\
  $\pi_B = \phi_B\sigma_{ABC}$, \\
  $\pi_C = \phi_C\sigma_{ABC}$.\\
  This makes the diagram for 
  $A\times B\times C$ commute by the commutativity of (2) and (3), 
  thereby forcing $\sigma_{ABC}=\sigma_{A(BC)}$ since the morphism
  $A\to A\times (B\times C)$ that makes (2) commute is unique.\\
  The proof for $(A\times B)\times C$ is entirely analogous, 
  giving us that $(A\times B)\times C \cong A\times (B\times C)$ by their
  satisfaction of the same universal property.
\end{proof}
\end{solution}
\begin{problem}{5.10}
  Push the envelope farther still, and define products for families (i.e. indexed sets) of objects of a category. Do these exist in \catname{Set}?
\end{problem}
\begin{solution}
  Let \catname{C} be a category and $S$ be a family of objects in \catname{C} indexed by a set $J$. 
  The product $P=\prod_J S$
  is an object of \catname{C} together with morphisms 
  $\pi_j \in \Homc{C}{P,S_j}$ for all $j\in J$ such that
  every $X\in\Objc{C}$ together with morphisms 
  $f_j\in\Homc{C}{X,S_j}$ admits a unique morphism
  $\sigma\in\Homc{C}{X,P}$ such that
  $f_j = \pi_j\sigma$.\\
  That these exist in \catname{Set} is equivalent to the Axiom of Choice.

\end{solution}
\end{document}
