%&pdflatex 
\documentclass[12pt]{article} 
\usepackage[margin=1in]{geometry} 
\usepackage{amsmath,amsthm,amssymb,amsfonts,tikz-cd} 

\newenvironment{problem}[2][Problem]{\begin{trivlist}
\item[\hskip \labelsep {\bfseries #1}\hskip \labelsep {\bfseries #2.}]}{\end{trivlist}}

\newenvironment{proposition}[1][Proposition]{\begin{trivlist}
\item[\hskip \labelsep {\bfseries #1.}]}{\end{trivlist}}

\newenvironment{definition}[1][Definition]{\begin{trivlist}
\item[\hskip \labelsep {\bfseries #1.}]}{\end{trivlist}}

\newcommand{\til}{\char`\~}
\newcommand{\bs}{\textbackslash} 
\newcommand{\nul}{\text{null}\ }
\newcommand{\range}{\text{range}\ }
\newcommand{\conj}[1]{\overline{#1}}

\newenvironment{solution}
  {\renewcommand\qedsymbol{$\blacksquare$}\begin{proof}[Solution]}
{\end{proof}}

\newenvironment{sproof}{
  \renewcommand\qedsymbol{$\square$}
  \begin{proof}
  }{
  \end{proof}
}

\begin{document}

\title{Linear Algebra Done Right Exercises\\ \large Chapter 4}
\author{David Melendez}
\maketitle

\begin{problem}{5}
  Suppose $m$ is a nonnegative integer, $z_1,\dots,z_{m+1}$ are distinct elements of $\mathbb{F}$, and
  $w_1,\dots,w_{m+1} \in \mathbb{F}$. Prove that there exists a unique polynomial 
  $p\in \mathcal{P}_m(\mathbb{F})$ such that
  \begin{equation*}
    p(z_j) = w_j
  \end{equation*}
  for $j = 1,\dots,m+1$.
\end{problem}
\begin{proof}
  Let $T \in \mathcal{L}(\mathcal{P}_m(\mathbb{F}), \mathbb{F}^X)$ where $X = \{z_1,\dots,z_{m+1}\}$. 
  Define $T$ by
  \begin{equation*}
    T(p)(x_j) = p(x_j)\\
  \end{equation*}
  It is easy to show that this transformation is linear.\\
  Now we compute the null space of $T$.
  Note that when $Tp=0$, we have that $p(x)=0$ for all $x\in X$, 
  and thus $p$ has at least $|X| = m+1$ distinct zeroes. 
  But since the degree of $p$ is at most $m$, this must mean that $p$ is the zero polynomial.
  Hence $T$ is injective and $\dim\nul T = 0$. 
  We then have
  \begin{align*}
    \dim\range T &= \dim \mathcal{P}_m(\mathbb{F}) - \dim\nul T\\
    &= m + 1 \\
    &\stackrel{\footnotemark}{=} \dim \mathbb{F}^X\text{.}
  \end{align*}
  \footnotetext{See the following Stack Exchange post:
  https://math.stackexchange.com/questions/2288812/finding-dimension-of-a-vector-space-v/2289241\#2289241}
  This shows that $\range T = \mathbb{F}$, and therefore that $T$ is an isomorphism between
  $\mathcal{P}_m(\mathbb{F})$ and $\mathbb{F}^X$. 
  Since $T^{-1}$ assigns a unique polynomial to each function 
  (i.e. set of ordered pairs distinct in the first slot) 
  as stated in the problem, this completes the proof.
\end{proof}
\begin{problem}{6}
  Suppose $p\in\mathcal{P}(\mathbb{C})$ has degree $m$. 
  Prove that $p$ has $m$ distinct zeros if and only if $p$ and its derivative $p'$ have no zeros in common.
\end{problem}
\begin{proof}
  We will prove the contrapositive.
  Suppose $z$ is a zero of both $p$ and $p'$.
  Note that
  \begin{align*}
    p(x) &= (x-z)q(x)\\
    p'(x) &= q(x) + (x-z)q'(x)
  \end{align*}
  We then have
  \begin{align*}
    0 &= p'(z) \\
    &= q(z)
  \end{align*}
  and hence $z$ is a zero of $q$. 
  This means that $(x-z)^2$ is a factor of $p$, and therefore it can have at most $m-1$ distinct zeros.
\end{proof}
\begin{problem}{7}
  Every polynomial of odd degree with real coefficients has a real zero.
\end{problem}
\begin{proof}
  According to 4.17, every polynomial with real coefficients can be factored into a product of linear
  and quadratic factors with real coefficients, with each linear factor corresponding to a real zero
  and each quadratic factor representing a pair of nonreal zeros. Since the polynomial in question is
  of odd degree, this factorization must have an odd number of linear factors, and thus at least one
  real zero.
\end{proof}
\begin{problem}{8}
  Define $T : \mathcal{P}(\mathbb{R}) \to \mathbb{R}^\mathbb{R}$ by
  \begin{equation*}
    Tp = \begin{cases}
      \dfrac{p-p(3)}{x-3} & \text{if  } x\neq 3 \\
      p'(3) & \text{if  } x = 3
    \end{cases}
  \end{equation*}
  Show that $Tp\in \mathcal{P}(\mathbb{R})$ for every polynomial $p\in\mathcal{P}(\mathbb{R})$ and that
  $T$ is a linear map.
\end{problem}
\begin{solution}
  Since $3$ is a zero of the polynomial $r(x) = p(x) - p(3)$, we have
  \begin{equation*}
    p(x) - p(3) = (x-3)q(x)
  \end{equation*}
  for some $q\in\mathcal{P}(\mathbb{R})$.
  Taking the derivative of both sides, we have
  \begin{equation*}
    p'(x) = (x-3)q'(x) + q(x)
  \end{equation*}
  and hence $p'(3) = q(3)$.
  Since $q(x) = (Tp)(x)$ for all $x\in\mathbb{R}$, we then have $Tp=q$.\\
\end{solution}
\begin{problem}{9}
  Suppose $p\in \mathcal{P}(\mathbb{C})$. Define $q:\mathbb{C}\to\mathbb{C}$ by
  \begin{equation*}
    q(z)=p(z)\conj{p(\conj{z})}
  \end{equation*}
  Prove that $q$ is a polynomial with real coefficients.
\end{problem}
\begin{solution}
  We will work with the corresponding factorizations.
  Let $m=\deg p$ and each $\zeta_k$ be a zero of $p$.
  We then have
  \begin{align*}
    q(z) &= p(z)\conj{p(\conj{z})}\\
    &= \prod_{k=0}^{m}(z-\zeta_k)\conj{\prod_{k=0}^{m}(\conj{z}-\zeta_k)}\\
    &= \prod_{k=0}^{m}(z-\zeta_k)\conj{(\conj{z}-\zeta_k)}\\
    &= \prod_{k=0}^{m}(z-\zeta_k)(z-\conj{\zeta_k})\\
    &= \prod_{k=0}^{m}(z^2 - 2(\text{Re } \zeta_k)z + |\zeta_k|^2z)
  \end{align*}
\end{solution}
\begin{problem}{10}
  Suppose $m$ is a nonnegative integer and $p\in\mathcal{P}_m(\mathbb{C})$ is such that there exist
  distinct real numbers $x_0,x_1,\dots,x_m$ such that $p(x_j)\in\mathbb{R}$ for $j=0,1,\dots,m$.
  Prove that all the coefficients of $p$ are real.
\end{problem}
\begin{solution}
  This follows directly from the uniqueness part of problem 5.
\end{solution}
\begin{problem}{11}
  
\end{problem}
\end{document}

